---
title: "AFEL R workshop"
author: "Finn"
date: 2021-03-08
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# Data wrangling (munging)
Code for a quick (~1 hour) overview of data wrangling in R. [Data wrangling is the process of transforming and mapping data from one "raw" data form into another format with the intent of making it more appropriate and valuable for a variety of downstream purposes such as analytics.][1]


Today we will cover:

* `table()`
* `unique()`
* `join()`
* `dplyr`
  * `mutate()`
  * `select()`
  * `filter()`
  * `summarise()`
  * `arrange()`
* pivoting from wide to long format (`pivot_longer`)

I am assuming you have R 4.0.0 or later installed. 

Install the packages we will use:
```{r eval = FALSE}
if (!require("dplyr")) install.packages("dplyr")
if (!require("tidyr")) install.packages("tidyr")
if (!require("nzffdr")) devtools::install_github("flee598/nzffdr")
```

Load packges:
```{r, message=FALSE}
library(nzffdr) # for example data 
library(dplyr)  # for data wrangling 
library(tidyr)  # for pivot_longer()
```


[`dplyr`](https://dplyr.tidyverse.org) is a part of the `tidyverse` concerned with the basics of data manipulation. 




[1]: https://en.wikipedia.org/wiki/Data_wrangling